How Slow do my Transient Typechecks Go?

### 2019-08-17 paper.tex
James edits to abstract, sec1

\begin{abstract}
One form of type checking used in gradually typed language is \emph{transient type checking}: when ever an object 'flows' through code with a type annotation, the object is dynamically checked to have the methods required by said annotation.

Though na\"ive implementations of this have a high runtime overhead, just-in-time compilation and optimisation in virtual machines can eliminate much of this overhead.  Unfortunately
the improvement is not uniform: while most type checks can be optimised
away, some will significantly decrease a program's
performance, and some may even increase it.  In this paper we investigate how benchmark results can be used to identify which type checks have the greatest effects on performance.
Using these techniques,
programmers can optimise programs by removing expensive type checks,
and VM engineers can identify new opportunities for compiler optimisation.

\sm{this is much to brief for me.
Transient typechecks is a very specific thing.
I believe, VM people need more context here.
It's a form of gradual typing, which requires comparably simple checks
at run time to check type violations.
Then, how do we investigate the issues? we should give a hint what the solution is. otherwise, the sentence in using these techniques doesn't make sense to me, because we didn't say anything about them.}
\end{abstract}

@@

There is a spectrum of different approaches to gradual typing
\cite{kafka18,bensurvey18icfp}.
At one end is ``pluggable types'' (as in Strongtalk \cite{strongtalk}) or ``erasure
semantics'' (as in 
%% Two important approaches are optional\citep{}
%% and gradual typing\citep{GiladPluggable2004,Siek2006,XXXSiek2015}.
%% These are applied to dynamic languages to reap the benefits of typing, but
%% unfortunately also have limitations.
%With optional or pluggable approaches such as 
%\citeurl{TypeScript,}{TypeScript}{Microsoft}{27 June
%2018}{https://www.typescriptlang.org/}
TypeScript
\citep{typeScriptECOOP}), here
%
all types are erased before execution, limiting the benefit of
types to statically typed portions of programs, whilst not providing 
any for of run time checking.  In the middle there are
``transient'' or ``type-tag'' checks (as in Reticulated Python)
% and ``concrete'' checks as in Thorn
, these offer first-order semantics, by checking
that an object's type constructor or supported methods match
any explicit type annotations that the object flows through.
%%%WRONG kjx
%, and checking only when data flows through those declarations
\cite{Siek2007,Bloom2009,concrete15,reticPython2014,Greenman2018}.
Reticulated Python also supports alternative ``monotonic'' semantics
which mutate objects by narrowing their concrete types when passed
into contexts with more specific types.
On the other end of the spectrum are behavioural
typechecks (as in Typed Racket \cite{typedScheme08,takikawa2012},
Gradualtalk \cite{gradualtalk14}),
and ``proxies'' (provided by Reticulated Python), these
support higher-order semantics, retaining
types until run 
time, and performing checks eagerly, thus detailed information
about type violations can be given as soon as possible via ``blame''
tracking \cite{blame2009,blameForAll2011}.
Finally, there is also ductile typing,
which dynamically interprets a static type system at run time
\cite{Ductile2011}.
%
%\mwh{Does a monotonic semantics fit on this spectrum somewhere?}
%\kjx{Not in the greenman/fellenstein spectrum, but I guess I'd put it
%  after concrete and before behavioural. Does it matter?}
%
% This said "any", but the ecoop paper specifically showed that 
Unfortunately, most gradual system with run-time semantics
(as opposed to type erasure as in TypeScript) currently
imposes a significant run-time performance overhead to provide such
semantics.  This has lead to a significant body of research,
on one hand developing techniques to optimise gradual typing 
\citep{Vitousek2017,Muehlboeck2017,Bauman2017,Richards2017,Greenman2018}
and on the other, developing techniques to evaluate the
resulting performance  \cite{Takikawa2016,Greenman2019jfp}.

>>>>

\begin{abstract}
One form of type checking used in gradually typed language is
\emph{transient type checking}: when ever an object `flows' through
code with a type annotation, the object is dynamically checked to
ensure it has the methods required by the annotation.

Although na\"ive implementations of transient type checks have a high runtime overhead, just-in-time compilation and optimisation in virtual machines can eliminate much of this overhead.  Unfortunately
the improvement is not uniform: while most type checks can be optimised
away, some will significantly decrease a program's
performance, and some may even increase it.  In this paper we investigate techniques that interpret benchmark results to identify which type checks have the greatest effects on performance.
Using these techniques,
programmers can optimise programs by removing expensive type checks,
and VM engineers can identify new opportunities for compiler optimisation.

\sm{this is much to brief for me.
Transient typechecks is a very specific thing.
I believe, VM people need more context here.
It's a form of gradual typing, which requires comparably simple checks
at run time to check type violations.
Then, how do we investigate the issues? we should give a hint what the solution is. otherwise, the sentence in using these techniques doesn't make sense to me, because we didn't say anything about them.}
\kjx{still like brief abstracts. at least I defined ``techniques''}
\end{abstract}

@@

There is a spectrum of different approaches to gradual typing
\cite{kafka18,bensurvey18icfp}.
At one end is ``pluggable types'' (as in Strongtalk \cite{strongtalk}) or ``erasure
semantics'' (as in 
%% Two important approaches are optional\citep{}
%% and gradual typing\citep{GiladPluggable2004,Siek2006,XXXSiek2015}.
%% These are applied to dynamic languages to reap the benefits of typing, but
%% unfortunately also have limitations.
%With optional or pluggable approaches such as 
%\citeurl{TypeScript,}{TypeScript}{Microsoft}{27 June
%2018}{https://www.typescriptlang.org/}
TypeScript
\citep{typeScriptECOOP}), where
%
all types are erased before execution, limiting the benefit of
types to statically typed portions of programs, while not providing 
any for of run time checking.  In the middle there are
``transient'' or ``type-tag'' checks (as in Reticulated Python),
% and ``concrete'' checks as in Thorn
which offer first-order semantics, by checking
that an object's type constructor or supported methods match
% any explicit type annotations
any static types
that the object flows through.
%%%WRONG kjx
%, and checking only when data flows through those declarations
\cite{Siek2007,Bloom2009,concrete15,reticPython2014,Greenman2018}.
Reticulated Python also supports alternative ``monotonic'' semantics
which mutate objects by narrowing their concrete types when passed
into contexts with more specific types.
On the other end of the spectrum are behavioural
typechecks (as in Typed Racket \cite{typedScheme08,takikawa2012},
Gradualtalk \cite{gradualtalk14}),
and ``proxies'' (provided by Reticulated Python), which
support higher-order semantics, retaining
types until run 
time, and performing checks eagerly, thus detailed information
about type violations can be given as soon as possible via ``blame''
tracking \cite{blame2009,blameForAll2011}.
Finally, there is also ductile typing,
which dynamically interprets a static type system at run time
\cite{Ductile2011}.
%
%\mwh{Does a monotonic semantics fit on this spectrum somewhere?}
%\kjx{Not in the greenman/fellenstein spectrum, but I guess I'd put it
%  after concrete and before behavioural. Does it matter?}
%
% This said "any", but the ecoop paper specifically showed that 
Unfortunately, most gradual systems with run-time semantics
(as opposed to type erasure as in TypeScript) typically
imposes significant run-time performance overheads to provide these
semantics.  This has lead to a significant body of research,
on one hand developing techniques to optimise gradual typing 
\citep{Vitousek2017,Muehlboeck2017,Bauman2017,Richards2017,Greenman2018}
and on the other, developing techniques to evaluate the
resulting performance  \cite{Takikawa2016,Greenman2019jfp}.
